Classless Java

### 2016-03-24 OOPSLA16/sections/CaseStudy.tex
spellcheck

Our refactoring applies the pattern presented in
Section~\ref{subsec:ep} to the existing Mumbler code base to improve
its modularity and extensibility. Using the refactored code base (560 SLOC) it is 
becomes possible to add new operations modularly, and to support
independent extensibility (677 SLOC). We add one more
operation \texttt{print} to both the original and the refactored code
base. In the original code base the pretty printer is added
non-modularly by modifying the existing code. In the refactored code
base the pretty printer is added modularly. 
Although the code in the refactored version is slightly increased (by 2.4\% SLOC), the
modularity is greatly increased, allowing for improved reusability and maintainability.
\end{comment}

@@

\vspace{0.5em}
{\small
\begin{tabular}{llll}
\hline
\rowcolor[HTML]{C0C0C0} 
Code               & SLOC & Code               & SLOC \\ \hline
original (\texttt{eval})         & 626  & original (\texttt{eval}+\texttt{print}) & 661  \\ 
\rowcolor[HTML]{C0C0C0}
refactored (\texttt{eval})         & 560  & refactored (\texttt{eval}+\texttt{print}) & 677  \\ \hline
\end{tabular}
}

>>>>

Our refactoring applies the pattern presented in
Section~\ref{subsec:ep} to the existing Mumbler code base to improve
its modularity and extensibility. Using the refactored code base (560 SLOC) it  
becomes possible to add new operations modularly, and to support
independent extensibility (677 SLOC). We add one more
operation \texttt{print} to both the original and the refactored code
base. In the original code base the pretty printer is added
non-modularly by modifying the existing code. In the refactored code
base the pretty printer is added modularly. 
Although the code in the refactored version is slightly increased (by 2.4\% SLOC), the
modularity is greatly increased, allowing for improved reusability and maintainability.
\end{comment}

@@

\vspace{0.5em}
{\small
\begin{tabular}{llll}
\hline
\rowcolor[HTML]{C0C0C0} 
Code               & SLOC & Code               & SLOC \\ \hline
original (\texttt{eval})         & 626  & original (\texttt{eval}+\texttt{print}) & 661  \\ 
refactored (\texttt{eval})         & 560  & refactored (\texttt{eval}+\texttt{print}) & 677  \\ \hline
\end{tabular}
}

### 2016-03-24 OOPSLA16/sections/Implementation.tex
space.
spellcheck

This process is facilitated by frameworks like Lombok~\cite{lombok}:
a Java library that aims at removing (or
reducing) Java boilerplate code via
annotations. \mixin was created using Lombok.
Figure~\ref{fig:lombok}~\cite{neildo2011blog} illustrates the flow of
the \mixin annotation.
First java source code is parsed into an abstract syntax tree (AST).
The AST is then captured by Lombok:
each annotated node is passed to
the corresponding (Eclipse or Javac) handler. The handler is
free to modify the information of the annotated node, or even inject new nodes (like methods, inner classes,
etc). Finally, the Java compiler works on the modified AST to generate bytecode.
%Note how during the compilation,
%no source code is changed, and no new source/temporary files are created.

@@

\paragraph{Direct modification of the AST}
Lombok alters the generation process of the class files,
by directly modifying the AST. Neither the source code is modified nor
new Java files are generated. Moreover, and probably more importantly,
Lombok supports generation of code \emph{inside} a class/interface,
which conventional Java annotation processors do not support. For
example, the standard \texttt{javax.annotation} processor, which is part of the
Java platform, only allows generation of \emph{new code}, and the
new code has to be written in \emph{new files}. Modification and/or
reinterpretation of existing code is not supported.

@@

\begin{figure*}[t]
\saveSpaceFig
\saveSpaceFig
\centering
\includegraphics[width=6in]{pdfs/screenshot4.png}
\caption{Generated methods shown in the Outline window of Eclipse and auto-completion.}
\label{fig:screenshot}
\saveSpaceFig
\end{figure*}

@@

\item{\bf Self coherence}: the generated code itself is well-typed. That is,
  type errors are not present in code the user has not written (for
  example \texttt{of} methods in Figure~\ref{fig:screenshot}).
In our case it means that either \mixin{} produces (in a controlled way) an
understandable error, or the interface can be successfully annotated and the generated code is well-typed.

>>>>

This process is facilitated by frameworks like Lombok~\cite{lombok}:
a Java library that aims at removing (or
reducing) Java boilerplate code via
annotations. \mixin was created using Lombok.
Figure~\ref{fig:lombok}~\cite{neildo2011blog} illustrates the flow of
the \mixin annotation.
First Java source code is parsed into an abstract syntax tree (AST).
The AST is then captured by Lombok:
each annotated node is passed to
the corresponding (Eclipse or Javac) handler. The handler is
free to modify the information of the annotated node, or even inject new nodes (like methods, inner classes,
etc). Finally, the Java compiler works on the modified AST to generate bytecode.
%Note how during the compilation,
%no source code is changed, and no new source/temporary files are created.

@@

\paragraph{Direct modification of the AST}
Lombok alters the generation process of the class files,
by directly modifying the AST. Neither the source code is modified nor
new Java files are generated. Moreover, and probably more importantly,
Lombok supports generation of code \emph{inside} a class/interface,
which conventional Java annotation processors do not support. For
example, the standard \texttt{javax.annotation} processor, which is part of the
Java platform, only allows generation of \emph{new code}, and the
new code has to be written in \emph{new files}. Modification and/or
reinterpretation of existing code are not supported.

@@

\begin{figure*}[t]
\saveSpaceFig
\saveSpaceFig
\centering
\includegraphics[width=5in]{pdfs/screenshot4.png}
\caption{Generated methods shown in the Outline window of Eclipse and auto-completion.}
\label{fig:screenshot}
\saveSpaceFig
\end{figure*}

@@

\item{\bf Self coherence}: the generated code itself is well-typed. That is,
  type errors are not present in code the user has not written (for
  example \texttt{of} methods in Figure~\ref{fig:screenshot}).
In our case, it means that either \mixin{} produces (in a controlled way) an
understandable error or the interface can be successfully annotated and the generated code is well-typed.

### 2016-03-24 OOPSLA16/sections/RelatedWork.tex
spellcheck

\paragraph{Models of \objectoriented programming}
There are traditionally two big styles of OOP: \classbased and
\prototypebased. The \classbased tradition started with Simula~\cite{Dahl66simula},
and is the dominant style of OOP today. Major languages such
as Java, C\#, C++ or Scala favour a \classbased OOP style. Very often
\classbased languages are statically typed and they have interfaces or
some other similar mechanisms such as traits. However, in contrast to
IB, interfaces or traits in CB are not meant to
replace classes, but rather, they complement classes. Normally
interfaces or traits define the signatures of the methods and
\emph{may} have some behaviour (such as in Java 8 interfaces or Scala
traits). However classes are still responsible for object
instantiation and handling state. \Prototypebased languages started
with Self~\cite{Ungar87self}, and are still widely used today via languages like
JavaScript. Frequently \prototypebased languages are dynamically
typed. IB is clearly closer to CB. However the
key difference is that classes are not needed in IB. Object interfaces recover the responsibility of object
instantiation, and abstract state operations provide a way to deal
with state. A primary motivation of IB is to improve
modularity and flexibility. Central to these goals are the
more expressive and automated forms of type-refinement, and support
for multiple interface inheritance.

@@

%\paragraph{Traits and Java's default methods}
Simplifying the mixins approach, traits~\cite{scharli03traits} draw a strong
line between units of reuse and object factories. Traits, as units of reusable
code, contain only methods as reusable functionality, ignoring state and state
initialization. Classes, as object factories, require functionality from
(multiple) traits. Java 8 interfaces are closely related to
traits: concrete method implementations are allowed (via the \textbf{default}
keyword) inside interfaces. The introduction of default methods opens the gate
for various flavors of multiple inheritance in Java. Traits offer an algebra
of composition operations like sum, alias and exclusion, providing explicit conflict
resolution. Former work~\cite{bono14} provides details on mimicking the trait
algebra through Java 8 interfaces.

>>>>

\paragraph{Models of \objectoriented programming}
There are traditionally two big styles of OOP: \classbased and
\prototypebased. The \classbased tradition started with Simula~\cite{Dahl66simula},
and is the dominant style of OOP today. Major languages such
as Java, C\#, C++ or Scala favour a \classbased OOP style. Very often
\classbased languages are statically typed and they have interfaces or
some other similar mechanisms such as traits. However, in contrast to
IB, interfaces or traits in CB are not meant to
replace classes, but rather, they complement classes. Normally
interfaces or traits define the signatures of the methods and
\emph{may} have some behaviour (such as in Java 8 interfaces or Scala
traits). However classes are still responsible for object
instantiation and handling state. \Prototypebased languages started
with Self~\cite{Ungar87self} and are still widely used today via languages like
JavaScript. Frequently \prototypebased languages are dynamically
typed. IB is clearly closer to CB. However the
key difference is that classes are not needed in IB. Object interfaces recover the responsibility of object
instantiation, and abstract state operations provide a way to deal
with state. A primary motivation of IB is to improve
modularity and flexibility. Central to these goals are the
more expressive and automated forms of type-refinement, and support
for multiple interface inheritance.

@@

%\paragraph{Traits and Java's default methods}
Simplifying the mixins approach, traits~\cite{scharli03traits} draw a strong
line between units of reuse and object factories. Traits, as units of reusable
code, contain only methods as reusable functionality, ignoring state and state
initialization. Classes, as object factories, require functionality from
(multiple) traits. Java 8 interfaces are closely related to
traits: concrete method implementations are allowed (via the \textbf{default}
keyword) inside interfaces. The introduction of default methods opens the gate
for various flavors of multiple inheritance in Java. Traits offer an algebra
of composition operations like sum, alias, and exclusion, providing explicit conflict
resolution. Former work~\cite{bono14} provides details on mimicking the trait
algebra through Java 8 interfaces.
